Mathematical Structure for Programming

### 2024-07-18 MS4P.tex
some puzzles about numbers

@@

>>>>

\newtheorem{puz}{Puzzle}[chapter]

@@

\section{puzzles}

\begin{puz}
  An `add-respects-or' test is a computable way to decide a true-or-false property of spuds-and-lemon numbers such that
  \begin{itemize}
  \item the test gives false for a $\lemon$;
    \item whenever you can split up any number $x$ as $y+z$ in any way, the test gives true for $x$ if and only if the test gives true either for $y$ or for $z$ or for both.
  \end{itemize}
  A `boring' test is one which always gives the same answer, no matter what number you test.
  \begin{enumerate}
  \item Find one boring add-respects-or test.
  \item Either find a different boring add-respects-or test, or explain why no such test exists.
  \item Find an add-respects-or test which is \emph{not} boring.
  \item Either find a different non-boring add-respects-or test, or explain why no such test exists.
  \end{enumerate}
(To show that two tests are different, you must be able to explain why there is some number where they disagree.)
\end{puz}

\begin{puz}
 Let us say that a number is `munky' if it is exactly divisible by three, `minky' if division by three leaves a remainder of one, and `manky' if division by three leaves a remainder of two. Explain how to test these properties of spuds-and-lemon numbers. You should be able to remove one spud at a time and test what is left. After all, you will eventually get to the lemon! Hint: it is easier to solve all three parts of this puzzle together, rather than taking the one at a time.
\end{puz}

\begin{puz}
Our method of addition replaced the left $\lemon$ by the right number. You could also consider computations which replace each $\spud$ in a number by the same something (which had better have a stick on the end). What can you compute in this way?
\end{puz}

\begin{puz}
  A `multiply-respects-or' test is a computable way to decide a true-or-false property of spuds-and-lemon numbers such that
  \begin{itemize}
  \item the test gives false for $\spud\lemon$;
  \item whenever you can split up any number $x$ as $y\times z$ in any way, the test gives true for $x$ if and only if the test gives true either for $y$ or for $z$ or for both.
  \end{itemize}
  \begin{enumerate}
  \item Are any of your `minky', `manky', `munky' tests multiply-respects-or? If so, which and why?
  \item In a black box, I have a multiply-respects-or test which is not boring. What answer does it give for the $\lemon$? No, you may not look in the box!
  \item Is divisibility by \emph{four} a multiply-respects-or test? Why?
  \item What are the non-boring multiply-respects-or tests? (Note: it's terribly unfair for me to ask this question at this stage, but we'll make sense of it eventually.)
  \end{enumerate}
\end{puz}
